A Parallel Algorithm for Solving Global Optimization Problems and Its Application for Tuning Hyperparameters of AI Methods

### 2025-04-14 2025/RuSCDays/RSD-2025.tex
Fig4 and Fig 5 are corrected

%(рус.) Пусть проведено k испытаний, их результаты получены и обработаны мастером, а на текущей итерации процессы-рабочие проводят испытания в точках множества S_k  = {x^(k+1),x^(k+2),…,x^(k+p) }, т.е. в этих точках испытания уже начались, но еще не завершены. Как только один из процессов-рабочих завершает проведение испытания в некоторой точке (например, x^(k+1)), он пересылает процессу-мастеру результаты испытания. Получив от рабочего процесса результаты испытания и обработав их мастер, в свою очередь, выбирает для него точку нового испытания x^(k+p+1) в соответствии со следующими правилами.
Suppose $k$ tests have been performed, the master has received and processed their results, and during the current iteration the worker processes perform trials at the points of the set $S_k={x^{k+1}, x^{k+2},..., x^{k+p}}$, i.e. the trials have already started at these points but have not yet been completed. As soon as one of the worker processes completes a trial at some point (e.g. $x^{k+1}$), it forwards the results of the trial to the master process. After receiving the trial results from the worker process and processing them, the master process in turn selects a new trial point $x^{k+p+1}$ for the worker process according to the following rules.

@@

%(рус.) Для решения поставленной задачи было выставлено ограничение на 1000 поисковых испытаний, число процессов P в параллельном алгоритме варьировалось от 1 до 20. В iOpt после 950 «глобальных» итераций глобального метода запускалось локальное уточнение с использованием Hooke-Jeeves method.
To solve the problem, a limit of 1000 search trials was set, the number of processes P in the parallel algorithm varied from 1 to 20. In iOpt, after 950 ``global'' iterations of the global method, local refinement using the Hooke-Jeeves method was carried out.

>>>>

%(рус.) Пусть проведено k испытаний, их результаты получены и обработаны мастером, а на текущей итерации процессы-рабочие проводят испытания в точках множества S_k  = {x^(k+1),x^(k+2),…,x^(k+p) }, т.е. в этих точках испытания уже начались, но еще не завершены. Как только один из процессов-рабочих завершает проведение испытания в некоторой точке (например, x^(k+1)), он пересылает процессу-мастеру результаты испытания. Получив от рабочего процесса результаты испытания и обработав их мастер, в свою очередь, выбирает для него точку нового испытания x^(k+p+1) в соответствии со следующими правилами.
Suppose $k$ trials have been performed, the master has received and processed their results, and during the current iteration the worker processes perform trials at the points of the set $S_k={x^{k+1}, x^{k+2},..., x^{k+p}}$, i.e. the trials have already started at these points but have not yet been completed. As soon as one of the worker processes completes a trial at some point (e.g. $x^{k+1}$), it forwards the results of the trial to the master process. After receiving the trial results from the worker process and processing them, the master process in turn selects a new trial point $x^{k+p+1}$ for the worker process according to the following rules.

@@

%(рус.) Для решения поставленной задачи было выставлено ограничение на 1000 поисковых испытаний, число процессов P в параллельном алгоритме варьировалось от 1 до 20. В iOpt после 950 «глобальных» итераций глобального метода запускалось локальное уточнение с использованием Hooke-Jeeves method.
To solve the problem, a limit of 1000 search trials was set, the number of processes P in the parallel algorithm varied from 1 to 20. In iOpt, after 950 ``global'' iterations of the global method, local refinement using the Hooke-Jeeves method was carried out.  %\cite{Himmelblau72}
